---
title: "STAT 5361 Interim Resport"
subtitle: "A Simple and Effective Inequality Measure"
author: 
   - Chamundeswari Koppisetti^[<chamundeswari.koppisetti@uconn.edu>] 
   - Shynggys Magzanov^[<shynggys.magzanov@uconn.edu>]
date: "19 11 2020"
documentclass: article
papersize: letter
fontsize: 11pt
biblio-style: asa
geometry: margin = 2.7cm
output: pdf_document

abstract: |
    We'll write it later
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

1) Examples of inequality curves and coefficients of inequality
for some common income distributions and compare values of I with G.
2) We introduce empirical versions of these concepts and investigate their inferential
properties, including robustness to outliers.
3) Applications to income data.

# 4. Application
## 4.1. Disposable Personal Income 

The data of Disposable Personal Income can be obtained at https://fred.stlouisfed.org/series/DSPI. 
```{r}
DSPI <- read.csv("C:\\Users\\Asus\\Downloads\\DSPI.csv")
plot(DSPI, type = "l", main = "Disposable Personal Income")
```

```{r}
QOR.ln <- function(u){
  # QOR function for the log-normal
  q.n.0 <- 1/dnorm(qnorm(u))
  q.n.1 <- qnorm(u)*q.n.0^2
  q.n.2 <- (1 + 2*qnorm(u)^2)*q.n.0^3
  1/(q.n.0^2 + 3*q.n.1 + q.n.2/q.n.0)
}

QOR.gld <- function(u, lambda = NULL){
  # QOR function for the GLD FKML paramaterization
  if(is.null(lambda)) lambda <- fit.fkml(x)$lambda
  l3 <- lambda[3]
  l4 <- lambda[4]
  (u^(l3 - 1) + (1 - u)^(l4 - 1))/(u^(l3 - 3)*(l3 - 2)*(l3 - 1) + 
                                     (1 - u)^(l4 - 3)*(l4 - 2)*(l4 - 1))
}

hatI <- function(x, J = 100, conf.level = 0.95, bw.correct = TRUE, QOR.FUN = QOR.ln, ...){
  
  n <- length(x)
  
  us <- ((1:J) - 0.5)/J
  Rs <- (xu2 <- quantile(x, us/2))/(x1u2 <- quantile(x, 1 - us/2))
  I <- sum(1 - Rs)/J
  
  if(!is.null(conf.level)){
    v <- c(us/2, 1 - us/2)
    qor <- QOR.FUN(v, ...)
    bw <- 15^(1/5)*abs(qor)^(2/5)/n^(1/5)
    if (bw.correct) bw[v <= bw] <- v[v <= bw]
    
    kernepach <- function(u) 3/4*(1 - u^2)*(abs(u) <= 1)
    m1 <- matrix(v, nrow = 2*J, ncol = n, byrow = FALSE)
    m2 <- matrix(1:n, nrow = 2*J, ncol = n, byrow = TRUE)
    
    consts <- kernepach((m1 - (m2 - 1)/n)*(1/bw))*(1/bw) - 
      kernepach((m1 - m2/n)*(1/bw))*(1/bw)
    
    x.sorted <- sort(x)
    q.hat <- c(consts%*%x.sorted)
    q.hat.1 <- q.hat[1:(length(q.hat)/2)]
    q.hat.2 <- q.hat[-(1:(length(q.hat)/2))]
    
    rc <- matrix(Rs, ncol = J, nrow = J, byrow = FALSE)
    
    covm <- ((1/x1u2)%*%t(1/x1u2))*(((us/2)%*%t(1 - us/2))*(q.hat.1%*%t(q.hat.1) + Rs%*%t(Rs)*(q.hat.2%*%t(q.hat.2))) -
                                      ((us/2)%*%t(us/2))*((q.hat.1%*%t(q.hat.2))*t(rc) + (q.hat.2%*%t(q.hat.1))*rc))/n
    
    sigma.p2 <- (us/2)*(1 - us/2)*q.hat.1^2
    sigma.q2 <- (1 - us/2)*(us/2)*q.hat.2^2
    sigma.pq <- (us/2)^2*q.hat.1*q.hat.2
    a0 <- sigma.p2/x1u2^2
    a1 <- -2*sigma.pq/x1u2^2
    a2 <- sigma.q2/x1u2^2
    Vs <- (a0 + a1*Rs + a2*Rs^2)/n
    
    V <- (sum(Vs) + 2*sum(covm[row(covm) < col(covm)]))/J^2
    SE <- sqrt(V)
    conf.int <- I + c(-1, 1)*qnorm(1 - (1 - conf.level)/2)*sqrt(V)
  } else{
    V <- NULL
    SE <- NULL
    conf.int <- NULL
  } 
  
  list(I = I, SE = SE, conf.int = conf.int)
}

hatG <- function(x, conf.level = 0.95){
  indices <- 1:(n <- length(x))
  ordered.x <- sort(x)
  sx <- sum(ordered.x*(indices - 1/2))
  mu.hat <- mean(x)
  Gv <- 2/mu.hat/n^2*sx - 1
  
  Z.hat <- -(Gv + 1)*ordered.x + (2*indices - 1)/n*ordered.x - 2/n*cumsum(ordered.x)
  Z.bar <- mean(Z.hat)
  
  V <- 1/n^2/mu.hat^2*sum((Z.hat - Z.bar)^2)
  conf.int <- Gv + c(-1, 1)*qnorm(1 - (1 - conf.level)/2)*sqrt(V)
  
  list(G = Gv, SE = sqrt(V), conf.int = conf.int)
}
```
```{r}
DSPI <- DSPI$DSPI
I <- hatI(DSPI)
I

G <- hatG(DSPI)
G

```

Firstly, $I$ and $G$ were calculated for the Disposable Personal Income Data. It appears that $\hat I=0.7561$, which is an estimate of $I$, with a standard error of $0.0106$. Additionally, 95% confidence interval of $\hat I$ was calculated to be $(0.7352, 0.7769)$. While $\hat G = 0.4872$ is significantly lower, with a standard error of $0.0095$ and respective 95% confidence interval of $(0.4686, 0.5059)$.

## 4.2. Gross Domestic Income 

The data on Gross Domestic Income can be found here: https://fred.stlouisfed.org/series/GDI. 

```{r}
GDI <- read.csv("C:\\Users\\Asus\\Downloads\\GDI.csv")
plot(GDI, type = "l", main = "Gross Domestic Income")
```
```{r}
GDI <- GDI$GDI
I_gdi <- hatI(GDI)
I_gdi

G_gdi <- hatG(GDI)
G_gdi
```

As a result, ratio of symmetric quantiles of the GDI is $\hat I=0.8204$ with a standard error $0.0162$ and confidence interval $(0.7886,0.8523)$. At the same time $\hat G=0.5499$ with a standard error of $0.0161$ and CI $(0.5182,0.5815)$. 

## 4.3. Earnings data of women

The data on Earnings of women in the US between 1979 and 2020 can be otbained from: https://fred.stlouisfed.org/series/LES1252882700Q. 
```{r}
LES <- read.csv("C:\\Users\\Asus\\Downloads\\LES.csv")
plot(LES, type = "l", main = "Median usual weekly nominal earnings for women")
```
```{r}
LES <- LES$LES
I_les <- hatI(LES)
I_les

G_les <- hatG(LES)
G_les
```

Results reveal that $\hat I=0.4666$ with a standard error $0.0183$ and CI $(0.4307,0.5025)$. While Gini coefficient is $\hat G=0.2199 \pm0.0094$ with CI $(0.2015,0.2384)$. 